---
title: "Churchill Seasonal Budgets"
author: "Dani Trangmoe"
date: "`r Sys.Date()`"
output: html_document
---

Seasonal Budgets

Use half-hourly data to calculate net co2 budgets by season Average rate per day

To separate graph for daily rate; use geom_violin


```{r, include=FALSE}
rm(list = ls())
Sys.setenv(TZ = "UTC")

library(data.table)
library(ggplot2)
library(lubridate)
library(tidyverse)
library(gridExtra)
library(patchwork) # for plot mapping
```


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = 'C:/Users/dtrangmoe/Documents/github/churchill_carbon_flux_paper/') #use this to set your directory
```


## Load Data
```{r}

df_all <- fread("./output_files/three_years_HH.csv")

```


## NEE

Plot 3 in Ameriflux poster

Finds the net CO2 exchange of each season as defined above

```{r}

seasonal_integrals_co2 <- function(df) {
  
  # Calculate integral (sum) for each season
  integrals <- df %>%
    group_by(season, full_year) %>%
    summarise(integral_value = sum(FC_F, na.rm = TRUE)) %>%
    ungroup() %>%
    arrange(season)
  
  
  # Changes units from  Micromoles of CO2/((m^2)s) to Grams of C/m^2, and rounds
    integrals$integral_value = 
      round(integrals$integral_value*60*30*(1/1000000)*12, 2)
  
  return(integrals)
}



# Full set

seasonal_integrals_co2 <- seasonal_integrals_co2(df_all)

seasonal_integrals_co2$season <- factor(seasonal_integrals_co2$season, levels = c("Fall Senescence 2022", "Winter 2023", "Snow Melt 2023", "Growing Season 2023", "Fall Senescence 2023", "Winter 2024", "Snow Melt 2024", "Growing Season 2024", "Fall Senescence 2024", "Winter 2025", "Snow Melt 2025", "Growing Season 2025"), ordered = TRUE)


seasonal_integrals_co2 <- seasonal_integrals_co2 %>%
  mutate(
    season_name = str_extract(season, "^[^0-9]+"), # Extract the season name
    season_year = str_extract(season, "\\d{4}")     # Extract the 4-digit year
  )




```


## Methane


```{r}

seasonal_integrals_ch4 <- function(df) {
  
  # Calculate integral (sum) for each season
  integrals <- df %>%
    group_by(season, full_year) %>%
    summarise(integral_value = sum(FCH4_F, na.rm = TRUE)) %>%
    ungroup() %>%
    arrange(season)
  
    
  # Changes units from nmol of CH4/(m^2s) to grams of C/m^2, and rounds
    integrals$integral_value = 
      round(integrals$integral_value*60*30*(1/1e9)*12, 2)
  
  return(integrals)
}


seasonal_integrals_ch4 <- seasonal_integrals_ch4(df_all)



seasonal_integrals_ch4$season <- factor(seasonal_integrals_ch4$season, levels = c("Fall Senescence 2022", "Winter 2023", "Snow Melt 2023", "Growing Season 2023", "Fall Senescence 2023", "Winter 2024", "Snow Melt 2024", "Growing Season 2024", "Fall Senescence 2024", "Winter 2025", "Snow Melt 2025", "Growing Season 2025"), ordered = TRUE)


seasonal_integrals_ch4 <- seasonal_integrals_ch4 %>%
  mutate(
    season_name = str_extract(season, "^[^0-9]+"), # Extract the season name
    season_year = str_extract(season, "\\d{4}")     # Extract the 4-digit year
  )


```

## Grouping

```{r}

#Join co2 and ch4
seasonal_integrals_net <- left_join(seasonal_integrals_co2, seasonal_integrals_ch4, by = c('season','season_year', 'season_name', 'full_year'), suffix = c('.co2','.ch4'))

#create net integral values
seasonal_integrals_net$integral_value.net <- seasonal_integrals_net$integral_value.co2 + seasonal_integrals_net$integral_value.ch4

unique(seasonal_integrals_net$season_name)

## Order seasons and years correctly
seasonal_integrals_net <- seasonal_integrals_net %>%
  mutate(season_name = factor(season_name,
                              levels = c("Fall Senescence ", "Winter ", "Snow Melt ", "Growing Season "), ordered = TRUE)) %>%
  mutate(full_year = factor(full_year,
                              levels = c("1", "2", "3"), ordered = TRUE))




```



## Plot

## NEE

```{r}
         
co2seasons <- ggplot(seasonal_integrals_net, aes(x = season_name, y = integral_value.co2, fill = full_year)) +
  theme_bw()+
theme(axis.text.x = element_text(size = 12), axis.text.y = element_text(size = 12), 
  axis.title = element_text(size = 14), legend.title = element_blank())+
  geom_hline(yintercept=0, colour="black")+geom_hline(yintercept=0, colour="black")+
  geom_col(position = position_dodge2(width = 0.9)) +
  labs(
    x = " ",
    y = expression("Net NEE (g C m"^-2*")"),
    title = expression("Net NEE by Season")
  ) +
  geom_label(aes(label = integral_value.co2, vjust = ifelse(integral_value.co2 >= 0, -1, 1.5)), colour = "black", fill ="white",
             position = position_dodge2(width = 0.9), size = 3)+
  scale_fill_manual(values = c("1" = "turquoise", "2" = "salmon", "3" = "navy"),
                    labels = c("1" = "2022/2023", "2" = "2023/2024", "3" = "2024/2025"))+
  scale_y_continuous(limits=c(-100, 20)) +
  scale_x_discrete(limits = c("Fall Senescence ", "Winter ", "Snow Melt ", "Growing Season "), labels = function(x) str_wrap(x, width = 10))

co2seasons

# png(filename = './co2seasons.png',width = 7.3,height = 4.1,units = 'in',res = 2000)
# co2seasons
# dev.off()
```


## Methane

```{r}

ch4seasons <- ggplot(seasonal_integrals_net, aes(x = season_name, y = integral_value.ch4,  fill = factor(full_year, order = TRUE))) +
  theme_bw()+
theme(axis.text.x = element_text(size = 12), axis.text.y = element_text(size = 12), 
  axis.title = element_text(size = 14), legend.title = element_blank())+
  geom_hline(yintercept=0, colour="black")+
  geom_col(position = position_dodge2(width = 0.9))+
  scale_fill_manual(values = c("1" = "turquoise", "2" = "salmon", "3" = "navy"),
                    labels = c("1" = "2022/2023", "2" = "2023/2024", "3" = "2024/2025"))+
  labs(
    x = "",
    y = expression("Net CH"[4] * " Flux (g C m"^-2*")"),
    title = expression("Net CH"[4] * " Flux by Season")
  ) +
  geom_label(aes(label = integral_value.ch4, vjust = -0.7), ,
            position = position_dodge2(width = 0.9), colour = "black", fill ="white", size = 3)+
  scale_y_continuous(limits=c(0, 6), expand = c(0,0))+
  scale_x_discrete(limits = c("Fall Senescence ", "Winter ", "Snow Melt ", "Growing Season "), labels = function(x) str_wrap(x, width = 10))

ch4seasons

# png(filename = './ch4seasons.png',width = 7.3,height = 4.1,units = 'in',res = 2000)
# ch4seasons
# dev.off()
 
```


Grid Plot

```{r}

png(filename = './png/seasongrid.png',width = 12,height = 9,units = 'in',res = 2000)
co2seasons + ch4seasons + plot_layout(guides = "collect", axes = "collect")
dev.off()


```



# GPP/RECO

## GPP

Finds the net GPP of each season

```{r}

seasonal_integrals_gpp <- function(df) {
  
  # Calculate integral (sum) for each season
  integrals <- df %>%
    group_by(season, full_year) %>%
    summarise(integral_value = sum(GPP_F, na.rm = TRUE)) %>%
    ungroup() %>%
    arrange(season)
  
  
  # Changes units from  Micromoles of CO2/((m^2)s) to Grams of C/m^2, and rounds
    integrals$integral_value = 
      round(integrals$integral_value*60*30*(1/1000000)*12, 2)
  
  return(integrals)
}



# Full set

seasonal_integrals_gpp <- seasonal_integrals_gpp(df_all)

seasonal_integrals_gpp$season <- factor(seasonal_integrals_gpp$season, levels = c("Fall Senescence 2022", "Winter 2023", "Snow Melt 2023", "Growing Season 2023", "Fall Senescence 2023", "Winter 2024", "Snow Melt 2024", "Growing Season 2024", "Fall Senescence 2024", "Winter 2025", "Snow Melt 2025", "Growing Season 2025"), ordered = TRUE)

#Filter to remove incomplete fall seasons
seasonal_integrals_gpp <- seasonal_integrals_gpp %>%
  filter(season %in% c("Fall Senescence 2022", "Winter 2023", "Snow Melt 2023", "Growing Season 2023", "Fall Senescence 2023", "Winter 2024", "Snow Melt 2024", "Growing Season 2024", "Fall Senescence 2024", "Winter 2025", "Snow Melt 2025", "Growing Season 2025"))


# Remove numeric year to group seasons
seasonal_integrals_gpp$season_name <- gsub(" [0-9]{4}", "", seasonal_integrals_gpp$season)

#change to negative for GPP

seasonal_integrals_gpp['integral_value'] <- -seasonal_integrals_gpp['integral_value']


```


## RECO


```{r}

seasonal_integrals_reco <- function(df) {
  
  # Calculate integral (sum) for each season
  integrals <- df %>%
    group_by(season, full_year) %>%
    summarise(integral_value = sum(RECO, na.rm = TRUE)) %>%
    ungroup() %>%
    arrange(season)
  
    
  # Changes units from nmol of reco/(m^2s) to grams of C/m^2, and rounds
    integrals$integral_value = 
      round(integrals$integral_value*60*30*(1/1000000)*12, 2)
  
  return(integrals)
}


seasonal_integrals_reco <- seasonal_integrals_reco(df_all)



seasonal_integrals_reco$season <- factor(seasonal_integrals_reco$season, levels = c("Fall Senescence 2022", "Winter 2023", "Snow Melt 2023", "Growing Season 2023", "Fall Senescence 2023", "Winter 2024", "Snow Melt 2024", "Growing Season 2024", "Fall Senescence 2024", "Winter 2025", "Snow Melt 2025", "Growing Season 2025"), ordered = TRUE)


#Filter to remove incomplete fall season
seasonal_integrals_reco <- seasonal_integrals_reco %>%
  filter(season %in% c("Fall Senescence 2022", "Winter 2023", "Snow Melt 2023", "Growing Season 2023", "Fall Senescence 2023", "Winter 2024", "Snow Melt 2024", "Growing Season 2024", "Fall Senescence 2024", "Winter 2025", "Snow Melt 2025", "Growing Season 2025"))


# Remove numeric year to group seasons
seasonal_integrals_reco$season_name <- gsub(" [0-9]{4}", "", seasonal_integrals_reco$season)


```

## Grouping

```{r}

seasonal_integrals_net <- left_join(seasonal_integrals_gpp, seasonal_integrals_reco, by = c('season','full_year', 'season_name'), suffix = c('.gpp','.reco'))

#create net integral values
seasonal_integrals_net$integral_value.net <- seasonal_integrals_net$integral_value.gpp + seasonal_integrals_net$integral_value.reco


## Order correctly

seasonal_integrals_net$season_name <- factor(seasonal_integrals_net$season_name, levels = c("Winter", "Snow Melt", "Growing Season", "Fall Senescence"), ordered = TRUE)



```




# Plot

## GPP

```{r}
         
gppseasons <- ggplot(seasonal_integrals_net, aes(x = season_name, y = integral_value.gpp, fill = factor(year, levels = c("2023", "2024")))) +
  theme_bw()+
theme(axis.text.x = element_text(size = 12), axis.text.y = element_text(size = 12), 
  axis.title = element_text(size = 14),legend.position = 'none')+
  geom_hline(yintercept=0, colour="black")+geom_hline(yintercept=0, colour="black")+
  geom_col(position = position_dodge2(width = 0.9)) +
  labs(
    x = " ",
    y = expression("Net GPP (g C m"^-2*")"),
    title = expression("Net GPP by Season")
  ) +
  geom_label(aes(label = integral_value.gpp, vjust = ifelse(integral_value.gpp >= 0, -1, 1.5)), colour = "black", fill ="white",
             position = position_dodge2(width = 0.9))+
  scale_fill_manual(values = c("2023" = "turquoise", "2024" = "salmon"))+
  scale_y_continuous(limits=c(-173, 0), expand = c(0,0)) +
  scale_x_discrete(labels = function(x) str_wrap(x, width = 10))

gppseasons

# png(filename = './gppseasons.png',width = 7.3,height = 4.1,units = 'in',res = 2000)
# gppseasons
# dev.off()
```


## RECO

```{r}

recoseasons <- ggplot(seasonal_integrals_net, aes(x = season_name, y = integral_value.reco,  fill = factor(year, levels = c("2023", "2024")))) +
  theme_bw()+
theme(axis.text.x = element_text(size = 12), axis.text.y = element_text(size = 12), 
  axis.title = element_text(size = 14),legend.position = 'none')+
  geom_hline(yintercept=0, colour="black")+
  geom_col(position = position_dodge2(width = 0.9))+
  scale_fill_manual(values = c("2023" = "turquoise", "2024" = "salmon"))+
  labs(
    x = "",
    y = expression("Net ER (g C m"^-2*")"),
    title = expression("Net Ecosystem Respiration by Season")
  ) +
  geom_label(aes(label = integral_value.reco, vjust = ifelse(integral_value.reco >= 0, -0.7, 1.3)), ,
            position = position_dodge2(width = 0.9), colour = "black", fill ="white")+
  scale_y_continuous(limits=c(0, 120), expand = c(0,0))+
  scale_x_discrete(labels = function(x) str_wrap(x, width = 10))

recoseasons

# png(filename = './ch4seasons.png',width = 7.3,height = 4.1,units = 'in',res = 2000)
# ch4seasons
# dev.off()
 
```


## Net Plot

```{r}



# netseasons <- ggplot(seasonal_integrals_net, aes(x = season_name, y = integral_value.net,  fill = factor(year, levels = c("2023", "2024")))) +
#   theme_bw()+
# theme(axis.text.x = element_text(size = 12), axis.text.y = element_text(size = 12), 
#   axis.title = element_text(size = 14),legend.position = 'none')+
#   geom_hline(yintercept=0, colour="black")+
#   geom_col(position = position_dodge2(width = 0.5))+
#   scale_fill_manual(values = c("2023" = "turquoise", "2024" = "salmon"), 
#                     guide = guide_legend(reverse = TRUE))+
#   labs(
#     x = "",
#     y = expression("Net Carbon Flux (g C m"^-2*")"),
#     title = expression("Net Carbon Flux by Season")
#   ) +
#   geom_text(aes(label = integral_value.net),
#             position = position_dodge2(width = 1),
#             vjust = ifelse(seasonal_integrals_net$integral_value.net >= 0, -1, 1.3),
#             colour = "black")+
#   scale_y_continuous(limits=c(-65, 30), expand = c(0,0))+
#   scale_x_discrete(labels = function(x) str_wrap(x, width = 10))
# 
# netseasons

# png(filename = './netseasons.png',width = 7.3,height = 4.1,units = 'in',res = 2000)
# netseasons
# dev.off()

```


Grid Plot

```{r}

grid.arrange(gppseasons, recoseasons, nrow = 1)

png(filename = './seasongrid_gpp_reco.png',width = 14,height = 7.2,units = 'in',res = 2000)
grid.arrange(gppseasons, recoseasons, nrow = 1)
dev.off()


```
