---
title: "Churchill Seasonal Budgets"
author: "Dani Trangmoe"
date: "`r Sys.Date()`"
output: html_document
---

Seasonal Budgets

Use half-hourly data to calculate net co2 budgets by season Average rate per day

To separate graph for daily rate; use geom_violin


```{r, include=FALSE}
rm(list = ls())
Sys.setenv(TZ = "UTC")

library(data.table)
library(ggplot2)
library(lubridate)
library(tidyverse)
library(gridExtra)
```



## Load Data
```{r}

df_23_24 <- fread("./bothyears_HH.csv")

```


## NEE

Plot 3 in Ameriflux poster

Finds the net CO2 exchange of each season as defined above

```{r}

seasonal_integrals_co2 <- function(df) {
  
  # Calculate integral (sum) for each season
  integrals <- df %>%
    group_by(season) %>%
    summarise(integral_value = sum(FC_F, na.rm = TRUE)) %>%
    ungroup() %>%
    arrange(season)
  
  
  # Changes units from  Micromoles of CO2/((m^2)s) to Grams of C/m^2, and rounds
    integrals$integral_value = 
      round(integrals$integral_value*60*30*(1/1000000)*12, 2)
  
  return(integrals)
}



# Full set

seasonal_integrals_co2 <- seasonal_integrals_co2(df_23_24)

seasonal_integrals_co2$season <- factor(seasonal_integrals_co2$season, levels = c("Fall Senescence 2022", "Winter 2023", "Snow Melt 2023", "Growing Season 2023", "Fall Senescence 2023", "Winter 2024", "Snow Melt 2024", "Growing Season 2024", "Fall Senescence 2024"), ordered = TRUE)

#Filter to remove incomplete fall seasons
filtered_data_co2 <- seasonal_integrals_co2 %>%
  filter(season %in% c("Winter 2023", "Snow Melt 2023", "Growing Season 2023", "Fall Senescence 2023", "Winter 2024", "Snow Melt 2024", "Growing Season 2024", "Fall Senescence 2024"))


# Remove numeric year to group seasons
filtered_data_co2$season_name <- gsub(" [0-9]{4}", "", filtered_data_co2$season)

#Add year factor
filtered_data_co2['year'] <- c(2023,2024,2023,2024,2023,2024,2023,2024)




```


## Methane


```{r}

seasonal_integrals_ch4 <- function(df) {
  
  # Calculate integral (sum) for each season
  integrals <- df %>%
    group_by(season) %>%
    summarise(integral_value = sum(FCH4_F, na.rm = TRUE)) %>%
    ungroup() %>%
    arrange(season)
  
    
  # Changes units from nmol of CH4/(m^2s) to grams of C/m^2, and rounds
    integrals$integral_value = 
      round(integrals$integral_value*60*30*(1/1000000000)*12, 2)
  
  return(integrals)
}


seasonal_integrals_ch4 <- seasonal_integrals_ch4(df_23_24)



seasonal_integrals_ch4$season <- factor(seasonal_integrals_ch4$season, levels = c("Fall Senescence 2022", "Winter 2023", "Snow Melt 2023", "Growing Season 2023", "Fall Senescence 2023", "Winter 2024", "Snow Melt 2024", "Growing Season 2024", "Fall Senescence 2024"), ordered = TRUE)


#Filter to remove incomplete fall season
filtered_data_ch4 <- seasonal_integrals_ch4 %>%
  filter(season %in% c("Winter 2023", "Snow Melt 2023", "Growing Season 2023", "Fall Senescence 2023", "Winter 2024", "Snow Melt 2024", "Growing Season 2024", "Fall Senescence 2024"))


# Remove numeric year to group seasons
filtered_data_ch4$season_name <- gsub(" [0-9]{4}", "", filtered_data_ch4$season)


#Add year factor
filtered_data_ch4['year'] <- c(2023,2024,2023,2024,2023,2024,2023,2024)



```

## Grouping

```{r}

filtered_data_net <- left_join(filtered_data_co2, filtered_data_ch4, by = c('season','year', 'season_name'), suffix = c('.co2','.ch4'))

#create net integral values
filtered_data_net$integral_value.net <- filtered_data_net$integral_value.co2 + filtered_data_net$integral_value.ch4


## Order correctly

filtered_data_net$season_name <- factor(filtered_data_net$season_name, levels = c("Winter", "Snow Melt", "Growing Season", "Fall Senescence"), ordered = TRUE)



```




## Plot

## NEE

```{r}
         
co2seasons <- ggplot(filtered_data_net, aes(x = season_name, y = integral_value.co2, fill = factor(year, levels = c("2023", "2024")))) +
  theme_bw()+
theme(axis.text.x = element_text(size = 12), axis.text.y = element_text(size = 12), 
  axis.title = element_text(size = 14),legend.position = 'none')+
  geom_hline(yintercept=0, colour="black")+geom_hline(yintercept=0, colour="black")+
  geom_col(position = position_dodge2(width = 0.9)) +
  labs(
    x = " ",
    y = expression("Net CO"[2] * " Flux (g C m"^-2*")"),
    title = expression("Net CO"[2] * " Flux by Season")
  ) +
  geom_label(aes(label = integral_value.co2, vjust = ifelse(integral_value.co2 >= 0, -1, 1.5)), colour = "black", fill ="white",
             position = position_dodge2(width = 0.9))+
  scale_fill_manual(values = c("2023" = "turquoise", "2024" = "salmon"))+
  scale_y_continuous(limits=c(-65, 23)) +
  scale_x_discrete(labels = function(x) str_wrap(x, width = 10))

co2seasons

# png(filename = './co2seasons.png',width = 7.3,height = 4.1,units = 'in',res = 2000)
# co2seasons
# dev.off()
```


## Methane

```{r}

ch4seasons <- ggplot(filtered_data_net, aes(x = season_name, y = integral_value.ch4,  fill = factor(year, levels = c("2023", "2024")))) +
  theme_bw()+
theme(axis.text.x = element_text(size = 12), axis.text.y = element_text(size = 12), 
  axis.title = element_text(size = 14),legend.position = 'none')+
  geom_hline(yintercept=0, colour="black")+
  geom_col(position = position_dodge2(width = 0.9))+
  scale_fill_manual(values = c("2023" = "turquoise", "2024" = "salmon"))+
  labs(
    x = "",
    y = expression("Net CH"[4] * " Flux (g C m"^-2*")"),
    title = expression("Net CH"[4] * " Flux by Season")
  ) +
  geom_label(aes(label = integral_value.ch4, vjust = ifelse(integral_value.ch4 >= 0, -0.7, 1.3)), ,
            position = position_dodge2(width = 0.9), colour = "black", fill ="white")+
  scale_y_continuous(limits=c(0, 6.2), expand = c(0,0))+
  scale_x_discrete(labels = function(x) str_wrap(x, width = 10))

ch4seasons

# png(filename = './ch4seasons.png',width = 7.3,height = 4.1,units = 'in',res = 2000)
# ch4seasons
# dev.off()
 
```


## Net Plot

```{r}



# netseasons <- ggplot(filtered_data_net, aes(x = season_name, y = integral_value.net,  fill = factor(year, levels = c("2023", "2024")))) +
#   theme_bw()+
# theme(axis.text.x = element_text(size = 12), axis.text.y = element_text(size = 12), 
#   axis.title = element_text(size = 14),legend.position = 'none')+
#   geom_hline(yintercept=0, colour="black")+
#   geom_col(position = position_dodge2(width = 0.5))+
#   scale_fill_manual(values = c("2023" = "turquoise", "2024" = "salmon"), 
#                     guide = guide_legend(reverse = TRUE))+
#   labs(
#     x = "",
#     y = expression("Net Carbon Flux (g C m"^-2*")"),
#     title = expression("Net Carbon Flux by Season")
#   ) +
#   geom_text(aes(label = integral_value.net),
#             position = position_dodge2(width = 1),
#             vjust = ifelse(filtered_data_net$integral_value.net >= 0, -1, 1.3),
#             colour = "black")+
#   scale_y_continuous(limits=c(-65, 30), expand = c(0,0))+
#   scale_x_discrete(labels = function(x) str_wrap(x, width = 10))
# 
# netseasons

# png(filename = './netseasons.png',width = 7.3,height = 4.1,units = 'in',res = 2000)
# netseasons
# dev.off()

```


Grid Plot

```{r}

grid.arrange(co2seasons, ch4seasons, nrow = 1)

png(filename = './seasongrid.png',width = 14,height = 7.2,units = 'in',res = 2000)
grid.arrange(co2seasons, ch4seasons, nrow = 1)
dev.off()


```



# GPP/RECO

## GPP

Finds the net GPP of each season

```{r}

seasonal_integrals_gpp <- function(df) {
  
  # Calculate integral (sum) for each season
  integrals <- df %>%
    group_by(season) %>%
    summarise(integral_value = sum(GPP_F, na.rm = TRUE)) %>%
    ungroup() %>%
    arrange(season)
  
  
  # Changes units from  Micromoles of CO2/((m^2)s) to Grams of C/m^2, and rounds
    integrals$integral_value = 
      round(integrals$integral_value*60*30*(1/1000000)*12, 2)
  
  return(integrals)
}



# Full set

seasonal_integrals_gpp <- seasonal_integrals_gpp(df_23_24)

seasonal_integrals_gpp$season <- factor(seasonal_integrals_gpp$season, levels = c("Fall Senescence 2022", "Winter 2023", "Snow Melt 2023", "Growing Season 2023", "Fall Senescence 2023", "Winter 2024", "Snow Melt 2024", "Growing Season 2024", "Fall Senescence 2024"), ordered = TRUE)

#Filter to remove incomplete fall seasons
filtered_data_gpp <- seasonal_integrals_gpp %>%
  filter(season %in% c("Winter 2023", "Snow Melt 2023", "Growing Season 2023", "Fall Senescence 2023", "Winter 2024", "Snow Melt 2024", "Growing Season 2024", "Fall Senescence 2024"))


# Remove numeric year to group seasons
filtered_data_gpp$season_name <- gsub(" [0-9]{4}", "", filtered_data_gpp$season)

#Add year factor
filtered_data_gpp['year'] <- c(2023,2024,2023,2024,2023,2024,2023,2024)

#change to negative for GPP

filtered_data_gpp['integral_value'] <- -filtered_data_gpp['integral_value']


```


## RECO


```{r}

seasonal_integrals_reco <- function(df) {
  
  # Calculate integral (sum) for each season
  integrals <- df %>%
    group_by(season) %>%
    summarise(integral_value = sum(RECO, na.rm = TRUE)) %>%
    ungroup() %>%
    arrange(season)
  
    
  # Changes units from nmol of reco/(m^2s) to grams of C/m^2, and rounds
    integrals$integral_value = 
      round(integrals$integral_value*60*30*(1/1000000)*12, 2)
  
  return(integrals)
}


seasonal_integrals_reco <- seasonal_integrals_reco(df_23_24)



seasonal_integrals_reco$season <- factor(seasonal_integrals_reco$season, levels = c("Fall Senescence 2022", "Winter 2023", "Snow Melt 2023", "Growing Season 2023", "Fall Senescence 2023", "Winter 2024", "Snow Melt 2024", "Growing Season 2024", "Fall Senescence 2024"), ordered = TRUE)


#Filter to remove incomplete fall season
filtered_data_reco <- seasonal_integrals_reco %>%
  filter(season %in% c("Winter 2023", "Snow Melt 2023", "Growing Season 2023", "Fall Senescence 2023", "Winter 2024", "Snow Melt 2024", "Growing Season 2024", "Fall Senescence 2024"))


# Remove numeric year to group seasons
filtered_data_reco$season_name <- gsub(" [0-9]{4}", "", filtered_data_reco$season)


#Add year factor
filtered_data_reco['year'] <- c(2023,2024,2023,2024,2023,2024,2023,2024)



```

## Grouping

```{r}

filtered_data_net <- left_join(filtered_data_gpp, filtered_data_reco, by = c('season','year', 'season_name'), suffix = c('.gpp','.reco'))

#create net integral values
filtered_data_net$integral_value.net <- filtered_data_net$integral_value.gpp + filtered_data_net$integral_value.reco


## Order correctly

filtered_data_net$season_name <- factor(filtered_data_net$season_name, levels = c("Winter", "Snow Melt", "Growing Season", "Fall Senescence"), ordered = TRUE)



```




# Plot

## GPP

```{r}
         
gppseasons <- ggplot(filtered_data_net, aes(x = season_name, y = integral_value.gpp, fill = factor(year, levels = c("2023", "2024")))) +
  theme_bw()+
theme(axis.text.x = element_text(size = 12), axis.text.y = element_text(size = 12), 
  axis.title = element_text(size = 14),legend.position = 'none')+
  geom_hline(yintercept=0, colour="black")+geom_hline(yintercept=0, colour="black")+
  geom_col(position = position_dodge2(width = 0.9)) +
  labs(
    x = " ",
    y = expression("Net GPP (g C m"^-2*")"),
    title = expression("Net GPP by Season")
  ) +
  geom_label(aes(label = integral_value.gpp, vjust = ifelse(integral_value.gpp >= 0, -1, 1.5)), colour = "black", fill ="white",
             position = position_dodge2(width = 0.9))+
  scale_fill_manual(values = c("2023" = "turquoise", "2024" = "salmon"))+
  scale_y_continuous(limits=c(-173, 0), expand = c(0,0)) +
  scale_x_discrete(labels = function(x) str_wrap(x, width = 10))

gppseasons

# png(filename = './gppseasons.png',width = 7.3,height = 4.1,units = 'in',res = 2000)
# gppseasons
# dev.off()
```


## RECO

```{r}

recoseasons <- ggplot(filtered_data_net, aes(x = season_name, y = integral_value.reco,  fill = factor(year, levels = c("2023", "2024")))) +
  theme_bw()+
theme(axis.text.x = element_text(size = 12), axis.text.y = element_text(size = 12), 
  axis.title = element_text(size = 14),legend.position = 'none')+
  geom_hline(yintercept=0, colour="black")+
  geom_col(position = position_dodge2(width = 0.9))+
  scale_fill_manual(values = c("2023" = "turquoise", "2024" = "salmon"))+
  labs(
    x = "",
    y = expression("Net ER (g C m"^-2*")"),
    title = expression("Net Ecosystem Respiration by Season")
  ) +
  geom_label(aes(label = integral_value.reco, vjust = ifelse(integral_value.reco >= 0, -0.7, 1.3)), ,
            position = position_dodge2(width = 0.9), colour = "black", fill ="white")+
  scale_y_continuous(limits=c(0, 120), expand = c(0,0))+
  scale_x_discrete(labels = function(x) str_wrap(x, width = 10))

recoseasons

# png(filename = './ch4seasons.png',width = 7.3,height = 4.1,units = 'in',res = 2000)
# ch4seasons
# dev.off()
 
```


## Net Plot

```{r}



# netseasons <- ggplot(filtered_data_net, aes(x = season_name, y = integral_value.net,  fill = factor(year, levels = c("2023", "2024")))) +
#   theme_bw()+
# theme(axis.text.x = element_text(size = 12), axis.text.y = element_text(size = 12), 
#   axis.title = element_text(size = 14),legend.position = 'none')+
#   geom_hline(yintercept=0, colour="black")+
#   geom_col(position = position_dodge2(width = 0.5))+
#   scale_fill_manual(values = c("2023" = "turquoise", "2024" = "salmon"), 
#                     guide = guide_legend(reverse = TRUE))+
#   labs(
#     x = "",
#     y = expression("Net Carbon Flux (g C m"^-2*")"),
#     title = expression("Net Carbon Flux by Season")
#   ) +
#   geom_text(aes(label = integral_value.net),
#             position = position_dodge2(width = 1),
#             vjust = ifelse(filtered_data_net$integral_value.net >= 0, -1, 1.3),
#             colour = "black")+
#   scale_y_continuous(limits=c(-65, 30), expand = c(0,0))+
#   scale_x_discrete(labels = function(x) str_wrap(x, width = 10))
# 
# netseasons

# png(filename = './netseasons.png',width = 7.3,height = 4.1,units = 'in',res = 2000)
# netseasons
# dev.off()

```


Grid Plot

```{r}

grid.arrange(gppseasons, recoseasons, nrow = 1)

png(filename = './seasongrid_gpp_reco.png',width = 14,height = 7.2,units = 'in',res = 2000)
grid.arrange(gppseasons, recoseasons, nrow = 1)
dev.off()


```
